taxize - taxonomic search and retrieval in R

### 2013-07-04 inst/doc/taxize_withcode_plos.Rnw
simplified classification example + some style

<<resolvenames, eval=TRUE, cache=TRUE>>=
temp <- gnr_resolve(names = c("Helianthos annus", "Homo saapiens"), returndf = TRUE)
temp[,-c(1,4)]
@

@@

<<tnrs, eval=TRUE,cache=TRUE>>=
mynames <- c("Helianthus annuus", "Pinus contort", "Poa anua", "Abis magnifica",
  	"Rosa california", "Festuca arundinace", "Sorbus occidentalos","Madia sateva")
tnrs(query = mynames)[,-c(5:7)]
@

@@

A number of data sources in taxize provide the capability to retrieve higher taxonomic names, but we will highlight two of the more useful ones: ITIS and NCBI. The principle in both is the same - first you need to get a numeric identifier for the queried species, then retrieve the higher taxonomy with this identifier. First, we'll explore the user of ITIS, by searching for two species, \emph{Abies procera} and \emph{Pinus contorta}.

<<itisclass, eval=TRUE,cache=TRUE>>=
specieslist <- c("Abies procera","Pinus contorta")
classification(get_tsn(specieslist, "sciname"))
@

It turns out both species are in the family Pinaceae. You can also get this type of information from the NCBI by doing \emph{classification(get\_uid(specieslist))}.

Instead of a full classification, you may only want a single name, say a family name for your species of interest. The function \emph{tax\_name} is built just for this purpose. And you can specify the data source you retrieve the taxonomic name from with the \emph{db} parameter, either ITIS or NCBI.

<<taxname, eval=TRUE,cache=TRUE>>=
tax_name(query="Helianthus annuus", get="family", db="itis")
tax_name(query="Helianthus annuus", get="family", db="ncbi")
@

\subsection*{Interactive name selection}
In the previous section we used the function \emph{get\_tsn} to get a classification for two plant species. Below are a few examples. When you run these examples in R, you are presented with a command prompt asking for the row that contains the name you would like back; that output is not printed below for brevity. In this example, the search term has many matches. The function returns a data.frame of the matches, and asks for the user to input what row number to accept.

<<interactive, eval=TRUE, echo=TRUE, cache=TRUE>>=
get_tsn(searchterm="Heliastes", searchtype="sciname")
@

@@

<<interactive_many, eval=TRUE, echo=TRUE, cache=TRUE>>=
splist <- c("annona cherimola", 'annona muricata', "quercus robur",
  	"shorea robusta", "pandanus patina", "oryza sativa", "durio zibethinus")
get_tsn(searchterm=splist, searchtype="sciname")
@

@@

<<interactive_na, eval=TRUE, echo=TRUE, cache=TRUE>>=
get_uid(sciname=c("Chironomus riparius", "aaa vva"))
@

@@

<<downstream, eval=TRUE, cache=TRUE>>=
col_downstream(name="Apis", downto="Species")[[1]]
@

@@

<<genbank_one, eval=TRUE, cache=TRUE>>=
out <- get_genes_avail(taxon_name="Umbra limi", seqrange = "1:2000", getrelated=FALSE)
@

@@

<<genbank_two, eval=TRUE, cache=TRUE>>=
out[grep("RAG1", out$genesavail, ignore.case=T),]
@

@@

<<aggregate_abd, eval=TRUE, cache=TRUE>>=
library(vegan)
data(dune.taxon)
dat <- dune.taxon
dat$abundance <- round(rlnorm(n=nrow(dat),meanlog=5,sdlog=2),0)
head( rankagg(data=dat, datacol="abundance", rank="Genus") )
head( rankagg(data=dat, "abundance", rank="Family", fxn="mean") )
@

>>>>

<<resolvenames, eval=TRUE, cache=TRUE>>=
temp <- gnr_resolve(names = c("Helianthos annus", "Homo saapiens"), returndf = TRUE)
temp[ , -c(1,4)]
@

@@

<<tnrs, eval=TRUE,cache=TRUE>>=
mynames <- c("Helianthus annuus", "Pinus contort", "Poa anua", "Abis magnifica",
  	"Rosa california", "Festuca arundinace", "Sorbus occidentalos","Madia sateva")
tnrs(query = mynames)[ , -c(5:7)]
@

@@

A number of data sources in taxize provide the capability to retrieve higher taxonomic names, but we will highlight two of the more useful ones: ITIS and NCBI. First, we'll search for two species, \emph{Abies procera} and \emph{Pinus contorta}.

<<itisclass, eval=TRUE,cache=TRUE>>=
specieslist <- c("Abies procera","Pinus contorta")
classification(specieslist, db = 'itis')
@

It turns out both species are in the family Pinaceae. You can also get this type of information from the NCBI by doing \emph{classification(specieslist, db = 'ncbi')}.

Instead of a full classification, you may only want a single name, say a family name for your species of interest. The function \emph{tax\_name} is built just for this purpose. As with the \emph{classification}-function you can specify the data source with the \emph{db} argument, either ITIS or NCBI.

<<taxname, eval=TRUE,cache=TRUE>>=
tax_name(query = "Helianthus annuus", get = "family", db = "itis")
tax_name(query = "Helianthus annuus", get = "family", db = "ncbi")
@

\subsection*{Interactive name selection}
As mentioned most databases use a numeric code to reference a species. A general workflow in taxize is: Retrieve Code for the queried species and then use this code to query more data/information.
Below are a few examples. When you run these examples in R, you are presented with a command prompt asking for the row that contains the name you would like back; that output is not printed below for brevity. In this example, the search term has many matches. The function returns a data.frame of the matches, and asks for the user to input what row number to accept.

<<interactive, eval=TRUE, echo=TRUE, cache=TRUE>>=
get_tsn(searchterm = "Heliastes", searchtype = "sciname")
@

@@

<<interactive_many, eval=TRUE, echo=TRUE, cache=TRUE>>=
splist <- c("annona cherimola", 'annona muricata', "quercus robur",
  	"shorea robusta", "pandanus patina", "oryza sativa", "durio zibethinus")
get_tsn(searchterm = splist, searchtype = "sciname")
@

@@

<<interactive_na, eval=TRUE, echo=TRUE, cache=TRUE>>=
get_uid(sciname = c("Chironomus riparius", "aaa vva"))
@

@@

<<downstream, eval=TRUE, cache=TRUE>>=
col_downstream(name = "Apis", downto = "Species")[[1]]
@

@@

<<genbank_one, eval=TRUE, cache=TRUE>>=
out <- get_genes_avail(taxon_name = "Umbra limi", seqrange = "1:2000", getrelated = FALSE)
@

@@

<<genbank_two, eval=TRUE, cache=TRUE>>=
out[grep("RAG1", out$genesavail, ignore.case = TRUE),]
@

@@

<<aggregate_abd, eval=TRUE, cache=TRUE>>=
library(vegan)
data(dune.taxon)
dat <- dune.taxon
dat$abundance <- round(rlnorm(n = nrow(dat), meanlog = 5, sdlog = 2), 0)
head( rankagg(data = dat, datacol = "abundance", rank = "Genus") )
head( rankagg(data = dat, "abundance", rank = "Family", fxn = "mean") )
@
